What does the crowd believe?

### 2016-08-11 presentations/01_CogSci_2016/slides.Rmd
CogSci slides makeover

---
title: "What does the crowd believe?"
subtitle: "A hierarchical approach to estimating subjective beliefs"
author: "Michael Franke, Fabian Dablander, Anthea Schöller (Tübingen)"
date : "Erin Bennett, Judith Degen, Michael Henry Tessler, Justine Kao & Noah D. Goodman (Stanford)"
runtime: shiny
output:
  ioslides_presentation:
    css: mistyle.css
    smaller: yes
    transition: faster
---
```{r setup, include=FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dev.args = list(bg = 'transparent'), fig.align='center')
require('ggplot2')
require('reshape2')
require('coda')
require('ggmcmc')
require('rjags')
require('runjags')
require('dplyr')
require('gridExtra')
require('rstan')
theme_set(theme_bw() + theme(plot.background=element_blank()) )
```

@@

## point of departure

@@

@@

- beliefs are latent, only act is directly observable

<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">INSERT A COOL PICTURE??</span>

## mundane beliefs may matter

<span style = "color:firebrick">sentence</span>

Joe eats many burgers.

@@

<span style = "color:firebrick">cardinal surprise reading</span>

Joe eats more burgers that we would expect of him.

<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>
<span style = "color:white"> &nbsp; </span>

<span style = "color:firebrick">INSERT A COOL PICTURE</span>

@@

- may not exist or may not be known

@@

- give-number task & inference of parameterized distribution 
    - e.g., Manski (2004); Tauber & Steyvers (2013)
    - <span style = "font-style: italic">scoring rules</span>
- binned histogram slider rating task (Kao et al., 2014)

## binned histogram slider ratings

@@

- relate subjects' answers, their beliefs & population-level aggregate

- data from multiple task types (within-subject)
- infer latent subjective & "population-level beliefs"
    - "population-level belief" => central tendency of subjective beliefs
    - think: "mean of a Gaussian hyperprior"

@@

## questions

@@

- relation: choice behavior, subjective & population-level beliefs

@@

<!--

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
</div>
-->

@@

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  red: averaged normal. slider ratings; black: mean posterior $Q_i$ with 95% HDIs
</div>

@@

<div style = "position:absolute; top: 620px; right:60px; color:darkblue;">
  black: mean posterior $Q_j$ with 95% HDIs; dark gray: mean posterior $P_{ij}$
</div>

@@

- subjective beliefs differ from population-level mean (good!)
- avrgd normlzd slider ratings reasonably approximate mean $Q_j$ (excellent!)

@@

avrgd normlzd slider ratings we would expect from the model and the posterior distribution over parameters is virtually indistinguishable from the observed

@@

some frequencies of number choices are suprising for the trained model (but: little data to go with; round or salient numbers may play a role)

@@

miserable failure to predict why it should be more likely that no marble sank than that one marble sank (alternative explanation: subjects revise beliefs, assume homogeneous "wonkiness" of marbles)

@@

- avrgd normlzd slider ratings appear to be practical and reliable measures

@@

- slider ratings closely track pop-level central tendency of individual beliefs

- hierarchical modeling of population-level beliefs is possible

- link functions for task choices seem reliable (enough)

@@

## link function: sliders

@@

## link function: numbers

@@

## link function: lightning

@@

>>>>

---
title: "What does the crowd believe?"
subtitle: "A hierarchical approach to estimating subjective beliefs"
author: "Michael Franke, Fabian Dablander, Anthea Schöller (Tübingen)"
date : "Erin Bennett, Judith Degen, MH Tessler, Justine Kao & Noah D. Goodman (Stanford)"
runtime: shiny
output:
  ioslides_presentation:
    css: mistyle.css
    smaller: yes
    transition: faster
---
```{r setup, include=FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, dev.args = list(bg = 'transparent'), fig.align='center')
require('ggplot2')
require('reshape2')
require('coda')
require('ggmcmc')
require('rjags')
require('runjags')
require('dplyr')
require('gridExtra')
require('rstan')
theme_set(theme_bw() + theme(plot.background=element_blank()) )
```

@@

## outline

<span style = "color:firebrick">Bayesian models of cognition</span>

- may require specifications of subjective beliefs

<span style = "color:firebrick">New experimental measure</span>

- binned histogram task

<span style = "color:firebrick">Bayesian data analysis</span>

- infer population-level and subjective beliefs from data

## beliefs are latent

@@

- beliefs are latent, only action is directly observable

@@

<div align = 'center'>
  <img src="pics/jones_umbrella.jpg" alt="JonesUmbrella" style="width: 600px;"/>
</div>

## focus: 1d continguous quantities

<div style = "float:left; width:45%;">
<span style = "color:firebrick">sentence</span>

Joe ate many burgers.

<span style = "color:white"> &nbsp; </span>
</div>
<div style = "float:right; width:45%;">
<span style = "color:firebrick">cardinal surprise reading</span>

Joe ate more than expected.

<span style = "color:white"> &nbsp; </span>
</div>

@@

<div style = "float:left; width:45%;">
```{r, echo = FALSE, fig.align='left', fig.width=3.5, fig.height=2.7}
bb = function(n,a,b) {
  sapply(0:n, function(k) choose(n,k) * beta(k+a,n-k+b) / beta(a,b))
}
theta = 0.8
MyDF<-data.frame(x = 0:25,
                 y = bb(25, 3, 10))
cutoff = max(which(cumsum(MyDF$y) <= theta)) - 1
p <-ggplot(MyDF, aes(x = x, y = y)) + geom_bar(stat = "identity", fill = "gray") +
  # geom_bar(data = filter(MyDF, x <= cutoff), stat = "identity", fill = "firebrick") +
  xlab("n") + ylab("") + 
  ggtitle("P(Joe ate n)")
show(p)
```  
</div>
<div style = "float:right; width:45%;">
```{r, echo = FALSE, fig.align='left', fig.width=3.5, fig.height=2.7}
MyDF2 = data.frame(x = 0:25,
                   y = cumsum(bb(25, 3, 10)))
cutoff = max(which(MyDF2$y <= theta)) - 1
p2 <-ggplot(MyDF2, aes(x = x, y = y)) + geom_bar(stat = "identity", fill = "gray") +
  geom_bar(data = filter(MyDF2, x <= cutoff), stat = "identity", fill = "firebrick") +
  xlab("n") + ylab("") +
  geom_hline(aes(yintercept=theta), color = "firebrick") +
  annotate("text", x = 3, y = 0.85, label = "theta = 0.8") +
  ggtitle("P(Joe ate n or less)")
show(p2)
```  
</div>

@@

- may not exist
- subjects may not be aware

@@

- give-number task & inference of parameterized distribution 
    - e.g., Manski (2004); Tauber & Steyvers (2013)
    - <span style = "font-style: italic">scoring rules</span> (Savage, 1971)
- iterated learning tasks (Lewandowsky, Griffiths, & Kalish, 2009)
- <span style = "color:firebrick">binned histogram task</span> (Kao et al., 2014)

<!--
## BH task

@@

-->

@@

@@

- take data from all three tasks
- infer latent subjective & "population-level beliefs"

<span style = "color:firebrick">main technical questions</span>

- how do individual-level beliefs relate to population level beliefs?
    - "population-level belief" => central tendency of subjective beliefs
    - think: "mean of a Gaussian hyperprior"
- how do subjective beliefs give rise to observable choices in the three tasks?

@@

@@

<div style = "position:absolute; top: 620px; right:60px;">
<span style = "color:firebrick">red</span>: BH task averages; <span style = "color:black">black</span>: mean posterior $Q_i$; <span style = "color:grey">grey</span>: 95% HDIs
</div>

@@

<div style = "position:absolute; top: 620px; right:60px;">
<span style = "color:black">black</span>: mean posterior $Q_i$; <span style = "color:grey">grey</span>: mean posterior $P_{ij}$
</div>

@@

- subjective beliefs $P_{ij}$ may differ from population-level mean (good!)

- BH task averages reasonably approximate mean $Q_j$ (excellent!)

@@

<div style = "position:absolute; top: 620px; right:60px;">
<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
</div>

@@

<div style = "position:absolute; top: 620px; right:60px;">
<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
</div>

@@

<div style = "position:absolute; top: 620px; right:60px;">
<span style = "color:firebrick">red</span>: observed data; <span style = "color:black">black</span>: mean posterior prediction with 95% HDIs
</div>

@@

- BH task averages are practical and reasonable approximations of the popuation-level central tendency of individual beliefs
- Bayesian data analysis is a great tool to extract information from multiple sources (e.g., data from different task types)

@@

- hierarchical modeling of population-level beliefs seems possible
- link functions for task choices seem reliable (enough)

@@

## link function: BH task

@@

## link function: GAN task

@@

## link function: PC task

@@

<div align = 'center'>
  <img src="//Users/micha/Desktop/data/svn/priors/models/04_Exp2_Manski_JAGS/plots/posterior_parameters.png" alt="postParameters" style="width: 650px;"/>
</div>
